Introductory Physics Labs: A Tale of Two Transformations

### 2020-08-18 manuscript/onlineLabPaper.tex
Updated introduction section for the draft.  Will work on next two sections tomorrow.  I will note that we are running a bit long and will either have to cut or re-work.  Including texcount info for completeness:
$ texcount onlineLabPaper.tex
File: onlineLabPaper.tex
Encoding: ascii
Words in text: 2095
Words in headers: 34
Words outside text (captions, etc.): 34
Number of headers: 9
Number of floats/tables/figures: 4
Number of math inlines: 0
Number of math displayed: 0
Subcounts:
  text+headers+captions (#headers/#floats/#inlines/#displayed)
  114+9+0 (2/0/0/0) _top_
  378+1+0 (1/0/0/0) Secti

\usepackage[colorlinks=true]{hyperref}
\usepackage{xcolor}
\usepackage{enumitem}
\setlist{nosep}

@@

\begin{abstract}
  A key problem facing physics departments, especially given the current global pandemic, is
  how can we engage students in our laboratory courses while maintaining appropriate social
  distancing and hygiene standards.  One solution to this problem is to move the labs to an
  online format.  But how can we engage our students in online labs with a curriculum that
  privileges science practices?  We have created an intro physics lab curriculum that engages
  students in science practices and are implementing it online.
\end{abstract}

@@

\section{Introduction}
\textcolor{blue}{SFW mostly writes this.}

Here we set up the rationale behind our initial transformation.  I see this as two-fold:
\begin{enumerate}
  \item We hadn't updated our labs in decades, and they utilized outdated technology and
  pedagogies.
  \item We wanted to give students an opportunity to delve into science practices.
\end{enumerate}
We have been seeking institutional (SFW's Teaching Grant) and national (NSF Award \#) resources
to support our initial transformation.

Then we've got a pandemic which means online labs were an immediate default. However, a lack of
online labs is also a barrier for DE students' completion of a degree.  So online lab curricula
would fill an institutional need.

@@

>>>>

\usepackage[colorlinks=true,urlcolor=blue,citecolor=blue]{hyperref}
\usepackage{xcolor}
\usepackage{enumitem}
\setlist{nosep}

@@

\begin{abstract}
  A key problem facing physics departments, especially given the current global pandemic, is
  how can we engage students in our laboratory courses while maintaining appropriate social
  distancing and hygiene standards.  One solution to this problem is to move the labs to an
  online format.  But how can we do this while \textit{also} engaging our students in online
  labs with a curriculum that privileges science practices?  We have created an intro physics
  lab curriculum that engages students in science practices and are implementing it online.
\end{abstract}

@@

\section{Introduction}
Scientific truth is discovered through the practice of doing science.  In fact, the National
Research Council in it's report on Discipline Based Education Research (DBER) states, ``Without
[scientific] investigative practices, there would be no new scientific and engineering
knowledge.''\cite{DBERreport} However, too often, laboratory courses do not meaningfully engage
students in scientific practices, instead reproducing classical experiments.\cite{PCAST12} This
is part of a long-standing tradition of calling for authentic engagement in scientific
practices which pre-dates Lorentz transformations.\cite{AAAS1881} Indeed, the AAAS stated our
goals quite succinctly, ``[Typical courses in t]he sciences\ldots are not made the means of
cultivating the observing powers, stimulating inquiry, exercising the judgment in weighing
evidence, nor of forming independent habits of thought.''\cite{AAAS1881}

The situation at our home institution in the Fall of 2016 very much mirrored the national
narrative.  The lab curriculum had been largely unchanged for about 20 years, and followed the
traditional cookbook lab paradigm.  Moreover, the equipment commonly used by students was
outdated -- triple beam balances and hand-drawn graphs were standard.  In this same year, as
part of an interdisciplinary team interested in engaging students in scientific practices as
well as assessing those scientific practices, we secured funding, both internally and from the
NSF, to change this status quo in physics, biology, and chemistry.  All three departments
transformed their labs using a pedagogical framework entitled, Argument Driven Inquiry
(ADI),\cite{Sampson2011,Walker2011,Walker2016} which has been specifically designed to engage
students in authentic science practices. We will focus on the physics transformation in this
paper.

We were satisfied with our progress when March 2020 rolled around, and with it, the mandatory
shut-down due to the global pandemic.  With students off campus, we had to abruptly shift our
delivery to an online format, and quickly realized that some of the activities that we utilized
in the face-to-face sections had to be adapted to online use.  Going forward, we've further
adapted what we are doing so that students can still participate in a hands-on laboratory
experience.  We hope these changes will be meaningful long-term as a lack of online labs is
also a barrier for DE students' completion of a degree.  So online lab curricula would fill an
institutional need.  In the next sections, we will discuss our institutional and instructional
contexts and describe each transformation in more detail.

@@

\begin{acknowledgments}
  Redacted for review.% This work has been supported by the NSF (Award \# 1725655).
\end{acknowledgments}
